A new tidy data structure to support exploration and modeling of temporal data

### 2019-01-28 main.Rmd
Section 3

(ref:tb-sub) A small subset of estimates of tuberculosis burden generated by World Health Organisation in 2011 and 2012, with 12 observations and 5 variables. The index refers to column `year`, the key to multiple columns: `country` and `gender`, and the measured variable to column `count`.

@@

The new data structure, tsibble, bridges the gap between raw temporal data and model inputs. Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs. Index, key and time interval are the three stone pillars to this new semantically structured temporal data.

@@

Time provides a contextual basis for temporal data. A variable representing time is essential for a tsibble, and is referred to as an "index". The "index" is an intact data column rather than a masked attribute, which makes time visible and accessible to users. It is highly advantageous when manipulating time. For example, one could easily extract time components, such as time of day and day of week, from the index to visualize seasonal effects of response variables. One could also join other data sources to the tsibble based on common time indexes. The accessibility of the tsibble index motivates data analysis towards transparency and human readability. When the "index" is available only as meta information (such as in the `ts` class), it created an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytic point of view.

A variable number of time representations can be spotted in the wild. A Date-time object, universally accepted across computing systems, is the most commonly used type for representing time. Date-time also typically associates with a time zone with adjustments such as summer time. This diversity and time zone is acknowledged and considered the building block for **tsibble**'s index.

@@

The "key" specification is the second essential ingredient for a tsibble. The "key" uniquely identifies observations that are recorded over time in a data table. It is similar to a primary key [@codd_relational_1970] defining each observation in a relational database. In the wide format in which multiple time series are often structured, the columns hold a series of values, so that the column implicitly serves as identification. In long format, columns are melted with names converted to "key" values. However, the "key" provides much more flexibility. It is not constrained to a single field, but can be composed from multiple fields. The identifying variables that the "key" is constituted from, remain the same as in the original table with no further tweaks.

@@

One of the cornerstones of time series data, and hence beneath a tsibble, is the time interval. This information plays a critical role in computing statistics (e.g. seasonal unit root tests) and building models (e.g. seasonal ARIMA). The principal divide is between regularly or irregularly spaced observations in time. A tsibble permits implicit missing time, making it difficult to distinguish regularity from the index. It relies on a user's specification by switching the `regular` argument off, when the data involves irregular intervals. This type of data can flow into event-based data modelling, but would need to be processed or regularized to fit models that expect time series.

For data indexed in regular time space, the time interval is automatically calculated, by first computing absolute differences of time indexes and then finding the greatest common divisor. This covers all conceivable cases. The implication is that all observations in a tsibble have one and only interval. Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each observation may different underlying data generating processes.

>>>>

(ref:tb-sub) A small subset of estimates of tuberculosis burden generated by World Health Organization in 2011 and 2012, with 12 observations and 5 variables. The index refers to column `year`, the key to multiple columns: `country` and `gender`, and the measured variable to column `count`.

@@

The new data structure, tsibble, bridges the gap between raw temporal data and model inputs. Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs. Index, key and time interval are the three stone pillars to this new semantically structured temporal data. Each is now described in more detail.

@@

Time provides a contextual basis for temporal data. A variable representing time is essential for a tsibble, and is referred to as an "index". The "index" is an intact data column rather than a masked attribute, which makes time visible and accessible to users. This is highly advantageous when manipulating time. For example, one could easily extract time components, such as time of day and day of week, from the index to visualize seasonal effects of response variables. One could also join other data sources to the tsibble based on common time indexes. The accessibility of the tsibble index motivates data analysis towards transparency and human readability. When the "index" is available only as meta information (such as in the `ts` class), it creates an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytic point of view.

A variable number of time representations can be spotted in the wild. A date-time object, universally accepted across computing systems, is the most commonly used type for representing time. Date-time also typically associates with a time zone including adjustments such as summer time. This diversity and time zone is acknowledged and accommodated by **tsibble**'s index.

@@

The "key" specification is the second essential ingredient for a tsibble. The "key" uniquely identifies observations that are recorded over time in a data table. It is similar to a primary key [@codd_relational_1970] defining each observation in a relational database. In the wide format in which multiple time series are often structured, the columns hold a series of values, so that the column implicitly serves as identification. In long format, columns are melted with names converted to "key" values. However, the "key" provides much more flexibility. It is not constrained to a single field, but can be composed from multiple fields. The identifying variables from which the "key" is constituted remain the same as in the original table with no further tweaks.

@@

One of the cornerstones of time series data, and hence beneath a tsibble, is the time interval. This information plays a critical role in computing statistics (e.g. seasonal unit root tests) and building models (e.g. seasonal ARIMA). The principal divide is between regularly or irregularly spaced observations in time. A tsibble permits implicit missing time, making it difficult to distinguish regularity from the index. It relies on a user's specification by switching the `regular` argument off, when the data involves irregular intervals. This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect time series.

For data indexed in regular time space, the time interval is automatically calculated, by first computing absolute differences of time indexes and then finding the greatest common divisor. This covers all conceivable cases, assuming that all observations in a tsibble have only one interval. Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each observation may have different underlying data generating processes.
